---
title: "Blain replication setup"
output: html_document
date: "2024-10-17"
---

# Description 

The code in all of these markdowns are designed to replicate the analyses from the Blain et al. (2021) paper. It is structured in the following way: 
  1. Calculate log(b) and log(d)
    a. Calculating all the statistics needed for the analyses.  
    b. Correlation between response bias in Block 1 and difference scores in bias from Block 1 to 2. 
  2. Effect of task manipulation. 
    a. Two-way repeated measures ANOVA: accuracy, block and type of stimulus as within subject factors.
    b. One-way repeated measure ANOVAs across each of the three blocks: 
      i. Discriminability 
      ii. Response bias 
    c. Followed by dependent sample t-tests
  3. Direct replication of depression and extraversion associations. 
    a. Model 1: repeated-measures ANOVA using [anhedonia scale] as a categorical variable with two levels. 
    b. Model 2: repeated-measures ANOVA using [anhedonia scale] as a continuous variable. 
    c. Correlations between each task performance variable and total anhedonia scores, and Big Five measures. 
    
    
# Explanation of .Rmd files: 

•	setup.Rmd: loading packages and data, and computing sdt statistics needed for analyses
•	task checks.Rmd: ANOVAs for accuracy, discriminability, and response bias, and reliability analyses for the task. 
•	descriptives.Rmd: descriptive statistics for the task and questionnaires, reliability for the questionnaires. 
•	models 1 & 2 - anhedonia.Rmd: Repeated measures ANOVAs and latent models for anhedonia (models 1 & 2). 
•	models 3 & 4 - extraversion.Rmd: Repeated measures ANOVAs and latent models for extraversion (models 3 & 4). 
• models 5-10.Rmd: all other latent variable models (including additional predictors - conscientiousness, and neuroticism; models 5-10). 
• computational model.Rmd: code for running the Huys et al. computational model. 

• making figures.Rmd: making all figures needed for the Method / Results sections. 
• additional analyses.Rmd: all analyses that are not directly done in the Blain paper, exploratory analyses and side quests. 
• testing PRT issues.Rmd: trying to figure out if there is an issue with the way the task has been coded. 

  

# Setup 

In this section, we load the data, and install/load all packages needed for the analyses. 

```{r 0.1 Setup, include=FALSE}

library(haven)
library(readr)
library(dplyr)

# Loading the packages 

library(lm.beta)
library(ez)
library(tidyverse)      # Data wrangling
library(psych)          # Used to undertake the EFA and related analyses 
library(GPArotation)    # Called by psych to run certain parts of the EFA 
library(EFA.dimensions) # Used for the RAWPAR function
library(lavaan)      # CFA/SEM
library(semPlot)     # For use of the `semPaths()` function to create the figure
library(semTools)    # Skew and kurtosis        
library(semhelpinghands)
library(dplyr)
library(naniar)
library(Hmisc)
library(corrplot)
library(ggpubr)
library(skimr)
library(MVN)
library(car)
library(aod)
library(knitr)
library(kableExtra)
library(data.table)
library(ggplot2)
library(stringr)
library(extrafont)
library(fontcm)
library(gridExtra)
library(tidyquant)
library(devtools)
library(ltm)
library(rstatix)
library(reshape2)
library(lsr)
library(ggraph)
library(igraph)
library(colorspace)
library(lme4)
library(lmerTest)


options(scipen = 999)

```

```{r 0.2 Loading data}

# Loading the questionnaire data 
questionnaire <- read_sav("study1_questionnaire_screened_NOV21.sav")
#View(questionnaire)

# Loading the task data 
task <- read_sav("study1_task_screened_NOV21.sav")
#View(task)


```

```{r 0.3 Adding HIT etc.}

# Adding stimulus type 
## 0 = lean 
## 1 = rich

task <- task %>%
  mutate(
    stimulus_type = case_when(
      
      # Group 1: expgroup = 1 (short = rich), responsekeyassignment = 1 (18 = short)
      values.expgroup == 1 & values.responsekeyassignment == 1 & trialcode == "shortmouth" & response == 18 ~ 1,  
      values.expgroup == 1 & values.responsekeyassignment == 1 & trialcode == "shortmouth" & response == 23 ~ 1,
      values.expgroup == 1 & values.responsekeyassignment == 1 & trialcode == "longmouth" & response == 23 ~ 0,  
      values.expgroup == 1 & values.responsekeyassignment == 1 & trialcode == "longmouth" & response == 18 ~ 0,
      
      # Group 2: expgroup = 1, responsekeyassignment = 2
      values.expgroup == 1 & values.responsekeyassignment == 2 & trialcode == "shortmouth" & response == 23 ~ 1,  
      values.expgroup == 1 & values.responsekeyassignment == 2 & trialcode == "shortmouth" & response == 18 ~ 1,
      values.expgroup == 1 & values.responsekeyassignment == 2 & trialcode == "longmouth" & response == 18 ~ 0,  
      values.expgroup == 1 & values.responsekeyassignment == 2 & trialcode == "longmouth" & response == 23 ~ 0, 
      
      # Group 3: expgroup = 2, responsekeyassignment = 2
      values.expgroup == 2 & values.responsekeyassignment == 2 & trialcode == "longmouth" & response == 18 ~ 1,  
      values.expgroup == 2 & values.responsekeyassignment == 2 & trialcode == "longmouth" & response == 23 ~ 1,
      values.expgroup == 2 & values.responsekeyassignment == 2 & trialcode == "shortmouth" & response == 23 ~ 0,  
      values.expgroup == 2 & values.responsekeyassignment == 2 & trialcode == "shortmouth" & response == 18 ~ 0, 
      
      # Group 4: expgroup = 2, responsekeyassignment = 1
      values.expgroup == 2 & values.responsekeyassignment == 1 & trialcode == "longmouth" & response == 23 ~ 1,  
      values.expgroup == 2 & values.responsekeyassignment == 1 & trialcode == "longmouth" & response == 18 ~ 1,
      values.expgroup == 2 & values.responsekeyassignment == 1 & trialcode == "shortmouth" & response == 18 ~ 0,  
      values.expgroup == 2 & values.responsekeyassignment == 1 & trialcode == "shortmouth" & response == 23 ~ 0,
      
      # Default case: If none of the conditions are met, set to NA
      TRUE ~ NA_real_  # Or set to 0 or another placeholder if you prefer
    )
  )


# Defining trial_type as "rich" and "lean"

task <- task %>%
  mutate(
    trial_type = case_when(
      
      # Group 1: expgroup = 1 (short = rich), responsekeyassignment = 1 (18 = short)
      values.expgroup == 1 & values.responsekeyassignment == 1 & trialcode == "shortmouth" ~ "rich",  
      values.expgroup == 1 & values.responsekeyassignment == 1 & trialcode == "longmouth"  ~ "lean",
      
      # Group 2: expgroup = 1, responsekeyassignment = 2
      values.expgroup == 1 & values.responsekeyassignment == 2 & trialcode == "shortmouth"  ~ "rich",  
      values.expgroup == 1 & values.responsekeyassignment == 2 & trialcode == "longmouth"  ~ "lean", 
      
      # Group 3: expgroup = 2, responsekeyassignment = 2
      values.expgroup == 2 & values.responsekeyassignment == 2 & trialcode == "longmouth" ~ "rich",  
      values.expgroup == 2 & values.responsekeyassignment == 2 & trialcode == "shortmouth" ~ "lean", 
      
      # Group 4: expgroup = 2, responsekeyassignment = 1
      values.expgroup == 2 & values.responsekeyassignment == 1 & trialcode == "longmouth" ~ "rich",  
      values.expgroup == 2 & values.responsekeyassignment == 1 & trialcode == "shortmouth" ~ "lean",
    )
  )


# Adding HIT, FA, MISS, CR

task <- task %>%
  mutate(
    # HIT: Correct response on a RICH trial
    hit = as.numeric(case_when(
      stimulus_type == 1 & correct == 1 ~ 1, 
      TRUE ~ 0  # Otherwise
    )),
    
    # FA: Incorrect response on a LEAN trial
    fa = as.numeric(case_when(
      stimulus_type == 0 & correct == 0 ~ 1,  # LEAN and incorrect
      TRUE ~ 0 
    )),
    
    # MISS: Incorrect response on a RICH trial
    miss = as.numeric(case_when(
      stimulus_type == 1 & correct == 0 ~ 1,  # RICH and incorrect
      TRUE ~ 0  # Otherwise
    )),
    
    # CR: Correct response on a LEAN trial
    cr = as.numeric(case_when(
      stimulus_type == 0 & correct == 1 ~ 1,  # LEAN and correct
      TRUE ~ 0  # Otherwise
    ))
  )





```

# Section 1

The code in this section is to replicate the analyses from the Blain et al. (2021) paper. It is structured in the following way: 
  1. Calculate log(b) and log(d)
    a. Calculating all the statistics needed for the analyses.  
    b. Correlation between response bias in Block 1 and difference scores in bias from Block 1 to 2. 
  2. Effect of task manipulation. 
    a. Two-way repeated measures ANOVA: accuracy, block and type of stimulus as within subject factors.
    b. One-way repeated measure ANOVAs across each of the three blocks: 
      i. Discriminability 
      ii. Response bias 
    c. Followed by dependent sample t-tests
  3. Direct replication of depression and extraversion associations. 
    a. Model 1: repeated-measures ANOVA using [anhedonia scale] as a categorical variable with two levels. 
    b. Model 2: repeated-measures ANOVA using [anhedonia scale] as a continuous variable. 
    c. Correlations between each task performance variable and total anhedonia scores, and Big Five measures. 
    

# 1. Calculate log(b) and log(d)

In this section, we are calculating all of the statistics needed to run the analyses in the Blain et al. (2021) paper. 

## 1.1 Calculating log(b) and log(d)

```{r 1.1.1 Ensuring data is numeric}


# Ensuring the data is numeric 

task <- task %>%
  mutate(hit = ifelse(hit == "1", 1, 0),
         fa = ifelse(fa == "1", 1, 0),
         miss = ifelse(miss == "1", 1, 0),
         cr = ifelse(cr == "1", 1, 0))
      

# Convert 'subject' to a factor
task$subject <- factor(task$subject)


```

```{r 1.1.2 blainstats: Creating homebase dataset with all statistics needed}

# Calculating the SDT statistics needed to replicate Blain's analyses 

blainstats <- task %>%
  group_by(subject, values.blockcount, values.responsekeyassignment, values.expgroup) %>%
  summarise(
    HIT = sum(hit), 
    FA = sum(fa), 
    MISS = sum(miss), 
    CR = sum(cr), 
    HR = HIT / (HIT + MISS),  
    FAR = FA / (FA + CR), 
    HR_adj = ifelse(HR == 0, 0.0001, ifelse(HR == 1, 0.9999, HR)),
    FAR_adj = ifelse(FAR == 0, 0.0001, ifelse(FAR == 1, 0.9999, FAR)),
    
    # calculating z-scores for the HIT and FALSE ALARM rates (and adjusting to account for 0 and 1)
    
    z_HR = qnorm(HR_adj),
    z_FAR = qnorm(FAR_adj),
    z_HR_adj = ifelse(z_HR == 0, 0.0001, z_HR),
    z_FAR_adj = ifelse(z_FAR == 0, 0.0001, z_FAR),
    
    # ordinate 
    
    o_HR = dnorm(z_HR_adj),
    o_FAR = dnorm(z_FAR_adj),
    
    # calculating SDT statistics 
    
    log_b = 0.5 * log((HIT + 0.5) * (FA + 0.5) / ((MISS + 0.5) * (CR + 0.5))),
    log_d = 0.5 * log((HIT + 0.5) * (CR + 0.5) / ((MISS + 0.5) * (FA + 0.5))),
    beta = o_HR / o_FAR,
    dprime = z_HR_adj - z_FAR_adj,
    .groups = 'drop'
  )%>%
  pivot_wider(
    names_from = values.blockcount, 
    values_from = c(HIT, FA, MISS, CR, HR, FAR, HR_adj, FAR_adj, z_HR, z_FAR, z_HR_adj, z_FAR_adj, o_HR, o_FAR, log_b, log_d, beta, dprime)
  ) %>%
  mutate(
    diffscore1_logb = log_b_2 - log_b_1, 
    diffscore2_logb = log_b_3 - log_b_1, 
    avlogb = rowMeans(dplyr::select(., starts_with("log_b_1"), starts_with("log_b_2"), starts_with("log_b_3")), na.rm = TRUE), 
    avlogb_baseline = rowMeans(dplyr::select(., starts_with("log_b_2"), starts_with("log_b_3")), na.rm = TRUE), 
    avbeta = rowMeans(dplyr::select(., starts_with("beta_1"), starts_with("beta_2"), starts_with("beta_3")), na.rm = TRUE),
    avlogd = rowMeans(dplyr::select(., starts_with("log_d_1"), starts_with("log_d_2"), starts_with("log_d_3")), na.rm = TRUE),
    avlogd_baseline = rowMeans(dplyr::select(., starts_with("log_d_2"), starts_with("log_d_3")), na.rm = TRUE), 
    avdprime = rowMeans(dplyr::select(., starts_with("dprime_1"), starts_with("dprime_2"), starts_with("dprime_3")), na.rm = TRUE),
  ) %>%
  rowwise() %>%
  mutate(
    total_hits = HIT_1 + HIT_2 + HIT_3, 
    total_cr = CR_1 + CR_2 + CR_3, 
    total_miss = MISS_1 + MISS_2 + MISS_3, 
    total_fa = FA_1 + FA_2 + FA_3,
    total_trials = total_hits + total_cr + total_miss + total_fa,
    accuracy = (total_hits + total_cr) / total_trials
  ) %>%
  ungroup() %>%
  mutate(
    bias_group = if_else(avlogb > 0, "Positive Bias", "Negative Bias")
  )

```

```{r 1.1.3 agg_trial: Creating aggregated dataset separated by block and trial_type}


agg_trial <- task %>%
  group_by(subject, values.blockcount, trial_type) %>%
  reframe(
    accuracy = case_when(
      trial_type == "lean" ~ sum(cr, na.rm = TRUE) / (sum(cr, na.rm = TRUE) + sum(fa, na.rm = TRUE)),
      trial_type == "rich" ~ sum(hit, na.rm = TRUE) / (sum(hit, na.rm = TRUE) + sum(miss, na.rm = TRUE)),
      TRUE ~ NA_real_  
    )
  ) %>%
  filter(!is.na(accuracy)) %>%
  distinct(subject, values.blockcount, trial_type, .keep_all = TRUE)

write.csv(agg_trial, file = "agg_trial.csv", row.names = FALSE)

```

```{r 1.1.4 agg_block: Creating aggregated dataset separated by block}


agg_block <- task %>%
  group_by(subject, values.blockcount, values.responsekeyassignment, values.expgroup) %>%
  summarise(
    HIT = sum(hit), 
    FA = sum(fa),
    MISS = sum(miss),
    CR = sum(cr), 
    correct = sum(correct),
    log_b = 0.5 * log((HIT + 0.5) * (FA + 0.5) / ((MISS + 0.5) * (CR + 0.5))),
    log_d = 0.5 * log((HIT + 0.5) * (CR + 0.5) / ((MISS + 0.5) * (FA + 0.5))),
    .groups = 'drop')



```

```{r 1.1.5 agg_merged: Merging agg_block with questionnaire data to be used in ANOVAs}

agg_merged <- agg_block %>%
  inner_join(questionnaire, by = c("subject" = "subject_q1"))

write.csv(agg_merged, file = "agg_merged.csv", row.names = FALSE)

``` 

```{r 1.1.6 merged_data: Merging the blainstats and questionnaire datasets}

merged_data <- blainstats %>%
  inner_join(questionnaire, by = c("subject" = "subject_q1"))

``` 
















